---
title: "Problem set 01"
date: "2024-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center")
source(here::here("R", "slide-things.R"))
```

## Task 1: Get up and running with R, RStudio, and the Tidyverse

The first thing we need to do is install the tools we'll be using. We'll follow the steps in [Example 01](/example/01-example.html) in class. 

## Task 2: Get introduced to R and RStudio

R is a programming language and environment for statistical computing. RStudio is an *integrated development environment*, or IDE for R. It is an application that acts as a kind of staging zone for the various related things you have to do when doing data analysis with R. It lets you type and save code in scripts or documents, run that code in R itself, see the results of what you have written, and manage various other tasks.

The best way to get started is to jump in directly. Read and follow along with Chapters 1 and 2 of Wickham et al's _R for Data Science_.


## Task 3: RStudio projects

For today, and for most of our problem sets, we will start with a zipped (i.e. compressed) file containing a project that has been mostly prepared in advance for you. 

```{r project-download, echo=FALSE, results="asis"}
project_buttons("/projects/01-problem-set")
```

Click the button above to download the project. Unzip it, and double-click the file named `01-problem-set.Rproj` to launch a new instance of RStudio. In RStudio, open the file named `analysis.qmd` and render it. Read the `qmd` file and compare it to its output.

## Task 4: Getting familiar with Quarto and RMarkdown

To ensure that the analysis and graphics you make are reproducible, you'll do the majority of your work in this class using **R Markdown** files. Watch this video, maybe more than once:

<div class="ratio ratio-16x9">
<iframe src="https://www.youtube.com/watch?v=_f3latmOhew" frameborder="0"  allow="autoplay; fullscreen" allowfullscreen></iframe>
</div>


