---
title: "Problem Set 07"
author: "<PUT YOUR NAME HERE>"
date: "`r Sys.Date()`"
output: html_document
---

```{r load-libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(socviz)
```

# Population Pyramids

First we will load some data and get it into the right shape:

```{r}
us <- read_csv(here("data", "uscenpops.csv")) |> 
  select(-pop) |> 
  pivot_longer(male:female, 
               names_to = "sex", 
               values_to = "pop")
```

Take a look at this dataset:

```{r}
us
```

The columns are as follows:

- `year`. : Calendar year
- `age`   : Years of age
- `sex`   : Recorded sex
- `pop`.  : Number of people of a given age and sex in a specific year.

# Questions

## How many males aged 50 were in the United States in 1920?

```{r}

```


## Draw a population pyramid for 1900

To get started, take a look at the notes for this week, where we draw a population pyramid using current population estimates. The one you draw here will be a little more straightforward because we don't have to fetch the data from the Census, and also because `age` will sort properly right away, as they are just numbers.

```{r}

```


## Compare the population pyramids for 1900, 1950, and 2000

Say a few words about how the US population seems to have changed over the course of the 20th century.
