---
title: "Example 05: Working with `dplyr`"
---


## Setup

```{r}

library(here)      # manage file paths
library(socviz)    # data and some useful functions
library(tidyverse) # your friend and mine
```

## Core `dplyr` verbs

```{r}

gss_sm  

```

### Select columns

```{r}
gss_sm |> 
  select(age, degree, bigregion, religion)

```

### Filter rows

```{r}
gss_sm |> 
  filter(age > 45)
```

```{r}
gss_sm |> 
  filter(childs > 4 & race == "White")
```

## Logically Group with `group_by()`

```{r}
gss_sm |> 
  group_by(bigregion)
```

### Summarize groups with `summarize()`

```{r}
gss_sm |> 
  group_by(bigregion) |>  #<<
  summarize(total = n()) 
```


### Multi-way groupings

```{r}

gss_sm |>  
  group_by(bigregion, religion) |> 
  summarize(total = n()) 
```


### Add columns with `mutate()`

```{r}
gss_sm |>  
  group_by(bigregion, religion) |> 
  summarize(total = n()) |> 
  mutate(freq = total / sum(total),
           pct = round((freq*100), 1))
```


## Tally and Count

- Do it yourself:

```{r}
gss_sm |> 
  group_by(bigregion, religion) |> #<<
  summarize(n = n()) #<<
```


- Use `tally()`:

```{r}

gss_sm |> 
  group_by(bigregion, religion) |> 
  tally() #<<
```

- Use `count()`:

```{r}
gss_sm |> 
  count(bigregion, religion) #<<
```

Pay attention to how grouping works in these summaries.

## Check your work


```{r}
rel_by_region <- gss_sm |> 
  count(bigregion, religion) |> 
  mutate(pct = round((n/sum(n))*100, 1)) 

rel_by_region
```

- Each region should sum to ~100

```{r}
rel_by_region |> 
  group_by(bigregion) |> 
  summarize(total = sum(pct)) 
```

- Grouping has caught us out. Try again.

```{r}

rel_by_region <- gss_sm |> 
  count(bigregion, religion) |> #<<
  mutate(pct = round((n/sum(n))*100, 1)) 

rel_by_region

```

## Summarize returns one tibble row per group

```{r}
gss_sm |> 
  group_by(bigregion) |> 
  tally()
```

When you have 2 or n-way groups the calculation is done from the inside out, on the innermost group. 

```{r}
# 4 regions, 6 religion = 24 groups
gss_sm |> 
  group_by(bigregion, religion) |> 
  tally()
```


## Summarize many variables

The inefficient way:

```{r}
organdata |>  
  group_by(consent_law, country)  |> 
    summarize(donors_mean= mean(donors, na.rm = TRUE),
              donors_sd = sd(donors, na.rm = TRUE),
              gdp_mean = mean(gdp, na.rm = TRUE),
              gdp_sd = sd(gdp, na.rm = TRUE),
              health_mean = mean(health, na.rm = TRUE),
              roads_mean = mean(roads, na.rm = TRUE),
              cerebvas_mean = mean(cerebvas, na.rm = TRUE))
```

## Use `across()` and `where()` instead

Better:

```{r}
organdata |> 
    group_by(consent_law, country) |>
      summarize(across(where(is.numeric),
                       list(mean = \(x) mean(x, na.rm = TRUE), 
                            sd = \(x) sd(x, na.rm = TRUE))))
```

The `\(x)` introduces an _anonymous function_ or _lambda_. The `x` means "the thing" or "the thing we're doing something to right now", and what follows it is some operation we perform on the thing.  

Optionally drop any remaning groups:

```{r}
organdata |> 
    group_by(consent_law, country) |>
      summarize(across(where(is.numeric),
                       list(mean = \(x) mean(x, na.rm = TRUE), 
                            sd = \(x) sd(x, na.rm = TRUE))),
                .groups = "drop")
```

- The `across()` function is used _inside_ `summarize()` and `mutate()` to do something across some subset of columns. 

- Inside `across()`, use `where()` to choose columns, and then apply a function to each of them. 

```{r}
organdata |>
  mutate(across(where(is.numeric), 
         round))
```

You can also use various "tidy selectors", like this:

```{r}
organdata |>
  mutate(across(starts_with("pop"), 
         round))
```


The function can be a named one, or you can write something yourself:

```{r}
organdata |>
  mutate(across(starts_with("pop"), 
         \(x) x / 100))
```

You can use `where()` with `select()` as well, when you are just subsetting by column but not yet doing anything across the columns:

```{r}
organdata |>
  select(where(is.character))
```

```{r}
organdata |>
  select(starts_with("gdp"))
```

```{r}
organdata |>
  select(contains("health"))
```


## Reminder: the `%in%` operator

This is a useful way to restrict selections of either columns, with `select()`, or especially rows, with (`filter`):

```{r}
organdata |> 
  filter(country %in% c("Ireland", "Italy", "Spain"))
```


## All this applies to `mutate()` as well

If you use a function like `mean()` or `sd()` or `n()` with `mutate()` instead of `summarize()` it will work too. The difference is that a column will be added with the value repeated for all group members. This can be useful when you want e.g. to make a denominator for some calculation later. Remember, `mutate()` adds or changes columns but never changes the number of rows in the table, whereas `summarize()` will usually output a table with fewer rows than the one you give it.

```{r}
## Country-year data, 238 rows altogether,
## with yearly data for 17 countries.
organdata |>
  select(country, donors)
```

```{r}
## Summarize gets you one row per country
organdata |>
  select(country, donors) |> 
  group_by(country) |> 
  summarize(donors_mean = mean(donors, na.rm = TRUE))

```


```{r}
## Mutate adds each country's donor mean 
## to the 238 observations
tmp <- organdata |>
  select(country, donors) |> 
  group_by(country) |> 
  mutate(donors_mean = mean(donors, na.rm = TRUE))

# First few rows of 238
head(tmp)

# Last few rows of 238
tail(tmp)
```


## Graph your summarized tables

```{r, fig.width=12, fig.height = 4}
gss_sm |> 
  group_by(bigregion, religion) |> 
  tally() |> 
  mutate(pct = round((n/sum(n))*100, 1)) |> 
  drop_na() |> 
  ggplot(mapping = aes(x = pct, 
                       y = reorder(religion, -pct), fill = religion)) + #<<
  geom_col() + #<<
    labs(x = "Percent", y = NULL) +
    guides(fill = "none") + 
    facet_wrap(~ bigregion, nrow = 1)

```


```{r}

rel_by_region <- gss_sm |> 
  group_by(bigregion, religion) |> 
  tally() |> 
  mutate(pct = round((n/sum(n))*100, 1)) |> 
  drop_na()


head(rel_by_region)
```


```{r}

p <- ggplot(data = rel_by_region, 
                mapping = aes(x = bigregion, 
                              y = pct, 
                              fill = religion))
p_out <- p + geom_col(position = "dodge") +
    labs(x = "Region",
         y = "Percent", 
         fill = "Religion") 

p_out
```


Experiment with facets:

```{r, fig.width=12, fig.height=4}
p <- ggplot(data = rel_by_region, 
                mapping = aes(x = pct, #<<
                              y = reorder(religion, -pct), #<<
                              fill = religion))
p_out_facet <- p + geom_col() +
  guides(fill = "none") + 
  facet_wrap(~ bigregion, nrow = 1) +
  labs(x = "Percent",
       y = NULL) 

p_out_facet
```



## Multi-way facets

```{r}
p <-  ggplot(data = gss_sm,
             mapping = aes(x = age, y = childs))

p + geom_point(alpha = 0.2) + 
  geom_smooth() +
  facet_wrap(~ race)
```



```{r}
p <-  ggplot(data = gss_sm,
             mapping = aes(x = age, y = childs))

p + geom_point(alpha = 0.2) + 
  geom_smooth() +
  facet_wrap(~ sex + race) #<<

```

```{r, fig.width=12, fig.height=4}
p <-  ggplot(data = gss_sm,
             mapping = aes(x = age, y = childs))

p + geom_point(alpha = 0.2) + 
  geom_smooth() +
  facet_wrap(~ sex + race, nrow = 1) #<<
```


### `facet_wrap()` vs `facet_grid()`

```{r}
p + geom_point(alpha = 0.2) + 
  geom_smooth() +
  facet_grid(sex ~ race) #<<

```


```{r, fig.width=12, fig.height=8}
p_out <- p + geom_point(alpha = 0.2) + 
  geom_smooth() +
  facet_grid(bigregion ~ race + sex) #<<

p_out
```


