---
title: "Example 12: Network Data"
---


## Setup

```{r}

library(here)       # manage file paths
library(socviz)     # data and some useful functions
library(tidyverse)  # your friend and mine

library(tidygraph) # tidy management of relational data
library(ggraph)   # geoms for drawing graphs

#remotes::install_github("kjhealy/kjhnet")
library(kjhnet)   # some network datasets
```


Sometimes, we have data that doesn't _look_ particularly relational at first glance, but we can make it so. 

```{r}
set_graph_style(family = "Myriad Pro SemiCondensed")

library(ggraph)
library(tidygraph)

set_graph_style(family = "Myriad Pro SemiCondensed")
```


## Warren's socjobs data

```{r}
socjobs
```

Let's clean it up a little:

```{r}
clean_dept_names <- function(x) {
    x <- str_replace(x, "California-", "")
    x <- str_replace(x, "SUNY-", "")
    x <- str_replace(x, "SUNY-", "")
    x <- str_replace(x, "Illinois-Chicago", "UIC")
    x <- str_replace(x, "U of Illinois", "UIUC")
    x <- str_replace(x, "U of Pennsylvania", "U Penn")
    x <- str_replace(x, "San Francisco", "UCSF")
    x <- str_replace(x, "North Carolina", "UNC")
    x <- str_replace(x, "N.C. State", "NC State")
    x
}

jobnet <- socjobs |> 
  # Texas isn't in the data properly
  filter(top25phd == "Yes", phd_dept != "Texas", job_dept != "Texas") |>
  select(phd_dept, job_dept) |>
  mutate(phd_dept = clean_dept_names(phd_dept),
         job_dept = clean_dept_names(job_dept)) |>
  group_by(phd_dept, job_dept) |>
  tally() 
```

At this point we have tallied jobs by PhD department and First Job Department:

```{r}
jobnet
```


Next we rename `n` to `weight` and make a scaled weight (in case we need it), and convert it to a network format:

```{r}

jobnet <- jobnet |> 
  select(from = phd_dept, to = job_dept, weight = n) |>
  mutate(scale_weight = scale(weight, center = FALSE)) |> 
  filter(weight > 0) |> 
  as_tbl_graph() 

jobnet

```

Now we can calculate centrality scores and draw a graph:

```{r, fig.height=10, fig.width=15}
jobnet |> 
  mutate(centrality = centrality_eigen(weights = weight)) |>
  ggraph(layout = "graphopt") + 
  geom_edge_fan(aes(alpha = weight),
                arrow = arrow(length = unit(3, "mm"), type = "closed"), 
                start_cap = circle(2, "mm"),
                end_cap = circle(8, "mm")) + 
  geom_node_label(aes(label = name)) + 
  scale_edge_alpha_continuous(name = "N Hires") + 
  labs(title = "New Assistant Professor Exchanges within the Top 25 Sociology Departments", 
       subtitle = "New Ph.D. hires, 1990-2017. Data show absolute number of within-network hires only.\nHires to and from outside the Top 25 are not shown. No adjustments are made for departmental or cohort sizes.", 
       caption = "Data: Warren (2019). Data exclude UT Austin.") + 
  theme_graph(base_family = "Myriad Pro SemiCondensed") + 
    theme(legend.position = "top") 
```
