---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| include = FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

workflow_graph <- readr::read_lines(here::here("merm.txt"))

```

## SOCIOL 232 / Visualizing Social Data / Spring 2024

## Kieran Healy, Duke University

This repo contains Course materials and website for [Visualizing Social Data](https://visualizingsociety.com), which [I](https://kieranhealy.org/about) am teaching at Duke University in the Spring of 2024. 

### Acknowledgements

Many thanks to [Andrew Heiss](https://www.andrewheiss.com) for, as usual, being ahead of the adoption curve with these tools and making several of his own courses [publicly available](https://www.andrewheiss.com/teaching/). Much of the back-end code and the `targets` pipeline used here has been adapted from Andrew's work. 
  

### Everything is a DAG now

This is the dependency graph for the site.

```{r, results = "asis", echo = FALSE}
cat("``` mermaid\n")
cat(workflow_graph, sep = "\n")
cat("\n```")  
```
