---
title: "11 --- Text as Data"
author: "Kieran Healy"
date: "`r Sys.Date()`"
output: kjhslides::kjh_local_slides_reader
editor_options: 
  chunk_output_type: console
---


```{r packages, include=FALSE}
library(flipbookr)
library(here)
library(tidyverse)
library(kjhslides)
```


```{r setup, include=FALSE}

kjh_register_tenso()
kjh_set_knitr_opts()
kjh_set_slide_theme()
kjh_set_xaringan_opts()






# Safe
```



class: center middle main-title section-title-1

# .kjh-green[Text] as .kjh-lblue[Data]

.class-info[

**Week 11**

.light[Kieran Healy<br>
Duke University, Spring 2023]

]

---

layout: true
class: title title-1

---

# Load the packages, as always

.SMALL[
```{r 07-iterating-on-data-2, message = TRUE}
library(here)      # manage file paths
library(socviz)    # data and some useful functions
library(tidyverse) # your friend and mine
```
]

---

# Specialty packages

.SMALL[
```{r 07-iterating-on-data-3, message = TRUE}
library(tidytext)    # Tools for analyzing text
library(gutenbergr)  # Get books from Project Gutenberg
library(janeaustenr) # Pre-organized dataset of Jane Austen's novels
```
]

---

# Toolkit

- This week's examples are mostly taken directly from Silge and Robinson (2017), _Text Mining with R_.

---

# Tidy text

```{r}
original_books <- austen_books() |>
  group_by(book) |>
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, 
                                     regex("^chapter [\\divxlc]",
                                           ignore_case = TRUE)))) |>
  ungroup()

tidy_books <- original_books |>
  unnest_tokens(word, text)

tidy_books
```

---

# "Stopwords"

- For many purposes (not always!) very common words like prepositions and articles are not interesting.

```{r}
data(stop_words)

stop_words

tidy_books <- tidy_books |>
  anti_join(stop_words)
```

---

# Stopwords removed

```{r}
tidy_books |>
  count(word, sort = TRUE) 
```


---

# Word frequency

```{r, fig.width=5, fig.height=4}
tidy_books |>
  count(word, sort = TRUE) |>
  filter(n > 600) |>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(n, word)) + geom_col() +
  labs(y = NULL)
```

---

# tf-idf

- The idea is to count the frequency of terms in a document, but decrease the weight of commonly used words and increase the weight for words that are not used very much in a corpus. 

---

# For example ...

```{r}
book_words <- austen_books() |>
  unnest_tokens(word, text) |>
  count(book, word, sort = TRUE)

total_words <- book_words |> 
  group_by(book) |> 
  summarize(total = sum(n))

book_words <- left_join(book_words, total_words)

book_words
```

---

# For example ...


```{r, fig.width=24, fig.height=4}
ggplot(book_words, aes(n/total, fill = book)) +
  geom_histogram(show.legend = FALSE) +
  xlim(NA, 0.0009) +
  facet_wrap(~book, nrow = 1, scales = "free_y")
```

---

# Zipf's Law

-  The frequency that a word appears is inversely proportional to its rank.

```{r}
freq_by_rank <- book_words |> 
  group_by(book) |> 
  mutate(rank = row_number(), 
         `term frequency` = n/total) |>
  ungroup()

freq_by_rank
```

---

# Zipf's Law

```{r, fig.height=6, fig.width=8}
freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = book)) + 
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) + 
  scale_x_log10() +
  scale_y_log10()
```


---

# n-grams

```{r}
austen_bigrams <- austen_books() |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  filter(!is.na(bigram))

austen_bigrams
```


---

# n-grams

```{r}
austen_bigrams |>
  count(bigram, sort = TRUE)
```

- Stopwords again

---

# n-grams

- Split the columns

```{r}
bigrams_separated <- austen_bigrams |>
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams_filtered <- bigrams_separated |>
  filter(!word1 %in% stop_words$word) |>
  filter(!word2 %in% stop_words$word)

# new bigram counts:
bigram_counts <- bigrams_filtered |> 
  count(word1, word2, sort = TRUE)

bigram_counts
```

---

# n-grams

```{r}
bigrams_united <- bigrams_filtered |>
  unite(bigram, word1, word2, sep = " ")

bigrams_united
```

- Now we have common bigrams without stopwords

---

# n-gram tf-idf

```{r}
bigram_tf_idf <- bigrams_united |>
  count(book, bigram) |>
  bind_tf_idf(bigram, book, n) |>
  arrange(desc(tf_idf))

bigram_tf_idf
```


---

# Plot them 

```{r}
out <- bigram_tf_idf |>
  arrange(desc(tf_idf)) |>
  group_by(book) |>
  slice_max(tf_idf, n = 10) |>
  ungroup() |>
  mutate(bigram = reorder(bigram, tf_idf)) |>
  ggplot(aes(tf_idf, bigram, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ book, nrow = 2, scales = "free") +
  labs(x = "tf-idf of bigram", y = NULL)
```


---

layout: false
class: center

```{r, fig.width=15, fig.height=6}
print(out)
```

---

layout: true
class: title title-1

---

# Sentiment Analysis

```{r}
raw_text <- gutenberg_download(4300) 

full_text <- raw_text |> 
  mutate(line=row_number()) |> 
  unnest_tokens(word,text)
```


---

# Sentiment Analysis

```{r}
full_text[30:51,] |> 
  print(n = Inf)
```


---

# Sentiment Analysis 

```{r}
tail(full_text, n = 20)
```

---

# Sentiment Analysis

```{r}
full_text |>
  anti_join(stop_words) |>
  filter(! str_detect(word, "'")) |>
  filter(! str_detect(word, "’")) |>
  count(word, sort = TRUE) |>
  top_n(20) |>
  mutate(word=reorder(word, n))
```


---

# Sentiment Analysis

```{r}
sent <- full_text |> 
  inner_join(get_sentiments("bing"), relationship = "many-to-many") |>
  count(sentiment, word, sort = TRUE) |>
  group_by(sentiment) |>
  top_n(10) |>
  ungroup() |>
  mutate(word=reorder(word,n)) 


sent

```


---

# Sentiment Analysis

```{r}
out <- sent  |> 
  ggplot(mapping = aes(x = n, 
                       y = word, 
                       fill=sentiment)) +
  geom_col() +
  guides(fill = "none") + 
  facet_wrap(~ sentiment, 
             scales = "free")

```


---

layout: false
class: center

```{r, fig.width=15, fig.height=5}
print(out)
```

---

layout: true
class: title title-1

---

# Sentiment Analysis

```{r}
out <- full_text |> 
  inner_join(get_sentiments("bing"), relationship = "many-to-many") |>
  count(index = line %/% 2000, sentiment) |>
  pivot_wider(names_from = sentiment, 
              values_from = n, values_fill = 0) |> 
  mutate(sentiment = positive - negative) 

out

```

---

# Sentiment Analysis

```{r, fig.height = 5, fig.width=15}
out |> 
  ggplot(mapping = aes(index, sentiment)) +
  geom_col(show.legend = FALSE)
```

---

# Pronouns


```{r}
pronouns <- raw_text |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  separate(bigram, c("word1", "word2"), sep=" ") |>
  filter(word1 %in% c("he", "she", "they")) |>
  filter(!word2 %in% stop_words$word, !str_detect(word2, "'")) |>
  count(word1, word2, sort=TRUE)

pronouns
```

---

# Pronouns

```{r}
out <- pronouns |> 
  group_by(word1) |> 
  top_n(15) |> 
  ggplot(mapping = aes(x=n, 
                       y=reorder(word2, n), 
                     fill=word1)) +
  geom_col() +
  guides(fill = "none") + 
  facet_wrap(~ word1, scales="free")
```

---

layout: false
class: center

```{r, fig.width=8, fig.height=8}
print(out)
```

