---
title: "Schedule"
sidebar: false
toc: false
page-layout: full
---

```{r packages-data, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)

withr::with_dir(here::here(), {
  schedule_nested <- targets::tar_read(schedule_page_data)
  targets::tar_load(schedule_ical_file)
})

base_url <- yaml::read_yaml(here::here("_quarto.yml"))$website$`site-url`
ical_url_absolute <- paste0(base_url, "/", schedule_ical_file)
```

::: {.schedule-page}

Here's a tentative outline for the semester. It's subject to change. There are up to three kinds of link each week. 

- [**Content**](/content/) ({{< fa book-open-reader >}}): These pages contain links to the readings and other material for each week. Read these **before** class. Lecture slide decks will also appear on these pages. 

- [**Example**](/example/) ({{< fa laptop-code >}}): These pages contain annotated R code and other supplementary information that you can use as a reference for your assignments and projects. It is also where any impromptu stuff we do in class will end up.  

- [**Assignment**](/assignment/) ({{< fa pen-ruler >}}): These pages contain the instructions for each problem set. The problem set problem set assiged in any given week is due the following _Tuesday_ by *6:00pm*. The midterm and final projects are due by *11:59 PM* on the day they're listed.

::: {.callout-tip}

## Subscribe to the class calendar

You can subscribe to this calendar URL in Outlook, Google Calendar, or Apple Calendar:

::: {.grid}

::: {.g-col-12 .g-col-md-2}
<p class="download-btn"><a class="btn btn-slide btn-sm" href="/`r schedule_ical_file`" role="button">{{< fa calendar-alt >}} Download</a></p>
:::

::: {.g-col-12 .g-col-md-10}
<p><input class="form-control" type="text" value="`r ical_url_absolute`" onclick="this.select()" readonly></p>
:::

:::

:::


```{r build-table, include=FALSE}
show_table <- function(group_id) {
  # Add a heading
  cat(as.character(paste("\n\n###", schedule_nested$group[[group_id]], "\n\n")))
  
  # Make the table
  tbl <- schedule_nested$data[[group_id]] %>% 
    select(-subgroup) %>% 
    kbl(escape = FALSE, align = "rlccc", table.attr = 'class="schedule-table"') %>% 
    kable_styling() %>% 
    column_spec(1, width = "20%", extra_css = "padding-right: 20px;") %>% 
    column_spec(2, width = "50%") %>% 
    column_spec(3:5, width = "10%") %>% 
    pack_rows(index = schedule_nested$subgroup_index[[group_id]],
              label_row_css = "border-bottom: 2px solid #000000;")
  
  cat(tbl)
}
```

```{r show-table, echo=FALSE, results="asis"}
walk(seq(1, nrow(schedule_nested)), ~show_table(.x)) 
```

:::
