---
title: "3. Ways of Seeing"
date: "2023-01-24"
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## Summary

This week we will take a step back from coding and begin to learn more about the principles behind the design of good visualizations. Many of these principles will lead to rules of thumb, guidelines, and techniques that we will end up applying in the code we write. But it is also important to situate these ideas a little more generally. So we start not just with information design or data in particular, but with some questions about the nature of visual representation in general. 


## Readings

These readings are available on the Canvas website, in the Files section. (Duke Login Required.)

- @berger72:_ways_seein Chapter 1 and Chapter 7. 
- @tufte83:_visual_displ_quant_infor Chapter 1.


## Video

In addition to reading Berger, I recommend watching the first of the episodes of the TV program that went with the book:

<iframe width="560" height="315" src="https://www.youtube.com/embed/0pDE4VX_9Kk" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>


## Slides

### Wednesday slides


```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/03-slides")
```

<div>
```{=html}
<iframe class="slide-deck" src="/slides/03-slides.html"></iframe>
```
</div>


