---
title: "Midterm"
author: "<PUT YOUR NAME HERE>"
date: "`r Sys.Date()`"
output: html_document
---

# Midterm Project

You will explore some data from the UK election in 2019. It is in the `data/` folder of this project in a CSV file called `ukelection2019.csv`

## Set up your work environment

First, load the tidyverse packages.

```{r}

```



## Read in the data

Next, read in the election data file. 

```{r}

```


## Display the first few rows of the file

```{r}

```


## Display a random sample of ten rows from the file

Use the `dplyr` function `slice_sample()` to do this. 

```{r}

```


## About the data

The `ukvote2019` data contains information on the the UK's general election in 2019. The _unit of observation_ is the _candidate_. Every row is a candidate who stood for election within a particular constituency, given by the `constituency` column. A candidate can only stand in one constituency. Candidates usually represent a political party, given by the `party_name` column. Candidates receive some number of votes (the `votes` column). The candidate with the most votes in a constituency wins. 

Some columns in the table refer specifically to the candidate. These include

- `party_name` : The party the candidate represents
- `candidate`, `fname`, and `lname` : The candidate's full name, first name, and last name
- `votes` : The number of votes cast for the candidate
- `vote_share_percent` : The percentage of votes cast received by the candidate. This is `(votes / total_votes_cast) * 100`, rounded to 1 decimal place. 


Some columns in the table refer to the constituency rather than the candidate, which means their values are the same for each candidate in that constituency. These include the name of the constituency, but also 

- `electorate` : The total number of people _eligible_ to vote in that constituency
- `total_votes_cast` : The number of people in the constituency who actually voted
- `turnout` : The proportion of eligble voters who actually voted (i.e., `total_votes_cast / electorate`)

Some columns in the table refer to the change in performance from the previous election. These include

- `vote_share_change` : Change in vote share for this party since the previous election, as a percentage.

## Questions

Explore the data by answering these questions:

## 1. Draw a histogram of the overall distribution of votes 

```{r}

```

Draw a histogram of the overall distribution of vote shares. 

```{r}

```

Briefly discuss what the histogram suggests. 

## 2. Draw a histogram of the overall distribution of turnout 

```{r}

```

How is it different from the histogram of votes? Why?

## 3. How many constituencies are there in the dataset?

```{r}

```


## 4. What percentage of all votes cast in the election went to the Conservative Party? 

```{r}

```

## 5. Using `geom_point()`, draw a graph of the total vote share for the five most popular parties.

```{r}

```


## 6. For the Conservative and Labour Parties only, was there a relationship between voter turnout and number of votes for the party?

```{r}

```

## 7. Draw a plot of the ten most common first names of candidates

```{r}

```


## 8. Optional bonus question: show some other feature of the data not already asked about

```{r}

```


