---
title: "Midterm"
author: "<PUT YOUR NAME HERE>"
date: "`r Sys.Date()`"
format: html
---

# Midterm Project

You will explore the movie data. It is in the `data/` folder of this project in a CSV file called `movies.csv`

## Set up your work environment

First, load the `tidyverse` package and the `here` package.

```{r}

```



## Read in the data

Next, read in the movie data file, putting in in an object called `movies`

```{r}

```


## Display the first few rows of the data

```{r}

```


## Display a random sample of ten rows from the data

Use the `dplyr` function `slice_sample()` to do this. 

```{r}

```


## About the data

The dataset was assembled and cleaned by Aaron Gullickson. It contains information on English-language movies produced in the US between 2000 and 2021; that are at least 80 minutes long and no longer than 3.5 hours long; that received at least 500 votes on the Internet Movie Database; that have an MPAA rating between G and R; and that made at least $100,000 domestically. 

The _unit of observation_ is an individual movie. That is, each row is a film. 

The columns or variables are:

- **year**: The calendar year of the film’s release.
- **runtime**: The length of the movie in minutes.
- **maturity_rating**: The movie’s [MPA maturity rating](https://en.wikipedia.org/wiki/Motion_Picture_Association_film_rating_system) (G, PG, PG-13, or R).
- **genre**: The genre of the film (one only).
- **box_office**: Gross domestic (US only) box office returns for the movie in millions of US dollars. Not adjusted for inflation.
- **rating_imdb**: This is average score (between 1 and 10) for a movie provided by IMDB users. 
- **metascore**: The movie's [metascore](https://www.metacritic.com/about-metascores) rating from [metacritic](https://www.metacritic.com/). The metascore is a curated weighted average of reviewer scores from a variety of sources.
- **awards**: The number of Oscar awards that this movie received.

## Questions

## 1. Draw a histogram of the distibution of movie run-time. 

```{r}

```

Say in a sentence or two what the distribution suggests to you. 

## 2. Draw a histogram of the distribution of box office receipts

```{r}

```


Say in a sentence or two what the distribution suggests to you. 

## 3. Is there a relationship between run-time and box office?

Use `geom_point()` and `geom_smooth()` to answer this question. Say in a sentence what you think the relationship is, and how strong you think it is.

```{r}

```

## 4. What do the distributions of IMDB rating and Metascore look like individually? 

```{r}

```


## 5. Is there a relationship between IMDB rating and Metascore? What sort?

```{r}

```

## 6. Draw a graph to show the distribution of box office returns by genre. Are some genres more likely than others to do well at the box office? Which ones?

```{r}

```

## 8. Draw a graph to show the number of movies made within each genre. 

```{r}

```

## 9. Draw a graph to show the number of awards received within each genre. 

```{r}

```

## 10. How is the number of awards received by a genre related to the number of movies made in that genre? Draw a graph showing the relationship. Label the genres on the graph. 

```{r}

```

## 11. Finally, some movies obviously do much better than others in terms of rating and box office. Draw a scatterplot of the relationship between metascore rating and box office for all movies, with labels showing the titles of the movies that score highly on both box office and metascore.

```{r}

```

