---
title: "6. Expanding Your Vocabulary"
date: "2023-02-15"
---

```{r slides, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "", "1", TRUE)

```

## Summary

This week we complete our initial tour of ggplot's capabilities by introducing ourselves to some more geoms and beginning to explore scale, guide, and theme functions.

## Reading

- @wickhamgrolemund:2017. [Chapter 5](https://r4ds.had.co.nz/transform.html), [Chapter 10](https://r4ds.had.co.nz/tibbles.html), and [Chapter 12](https://r4ds.had.co.nz/tidy-data.html).

## Slides

The slide deck for this week is available here. 

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/06-slides")
```

<div>
```{=html}
<iframe class="slide-deck" src="/slides/06-slides.html"></iframe>
```
</div>


:::{.callout-tip}
**Tip**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::
