---
title: "Problem Set 12"
author: "<PUT YOUR NAME HERE>"
date: "`r Sys.Date()`"
format: html
---

```{r}
#| label = "load-libraries-data",
#| warning = FALSE,
#| message = FALSE
library(tidyverse)
library(here)
library(socviz)

## Make sure these are installed
library(tidygraph)
library(ggraph)
```

# Star Wars Network

First we will load some data, collected by [Evelina Gabasova](https://zenodo.org/record/1411479#.ZDActC_MKgQ). Unlike other problem sets, these data are already an r object called `starwars`:

```{r}
load(here("data", "starwars.rda"))
```

The data are in tidygraph's linked two-table format, one for the nodes and one for the edges:

```{r}
starwars
```


# Graph the interaction network

```{r}
starwars |> 
  ggraph(layout = "fr") + 
  geom_edge_link(color="gray") + 
  geom_node_point(colour="lightblue", size=3) + 
  geom_node_text(aes(label=name), repel=T) +  
  theme_graph()
```

# Q1 What happens if you draw this graph more than once?

# Q2 Experiment with some alternative layouts. 

- Try `"fr"`, `"circle"`, `"randomly"`, and `"grid"` as arguments to `ggraph(layout = )`

- Which seem more or less useful? What seems to be changing in these layouts and what is staying the same?

# Q3 Calculate some graph statistics

- Like we did for the _Iliad_, calculate the _degree_, _betweenness_, and _alpha_ centrality for the characters and graph each result. Who is the most central character by each of these measures? Which results seem more or less plausible to you?

_Hint:_ To operate on the table of characters, use `activate_nodes()`. Because ties are not directed in this graph, when calculating alpha centrality you won't need to activate the edges or use `reroute()`. 

