---
title: "Problem Set 01"
author: "<YOUR NAME HERE>"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
#| label = "setup",
#| include = FALSE
knitr::opts_chunk$set(echo = TRUE)
```

## Quarto and R Markdown

This is an Quarto document. It is a structured plain-text document format that allows you to mix text and code in a way that can be turned into HTML, PDF, and MS Word documents. For more details on using Quarto see <https://quarto.org/docs/get-started/hello/rstudio.html>.

When you click the **Render** button in RStudio a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
#| label = "addition"
1 + 1
```

## Installing Support Packages

We will only need to run this chunk _once_, so this code chunk has a flag or option not to run when the document is knitted.

```{r}
#| label = "packages",
#| eval = FALSE
course_packages <- c("tidyverse", "babynames", "broom",
    "drat", "gapminder", "here", "janitor", "naniar", 
    "palmerpenguins", "skimr", "slider", "socviz",
    "usethis", "visdat","reprex")

install.packages(course_packages, repos = "http://cran.rstudio.com")

data_packages <- c("covdata", "congress", "nycdogs", 
                   "ukelection2019", "uscenpops")

drat::addRepo("kjhealy")
install.packages(data_packages)
```


## The Tidyverse Packages

In this next chunk we will load the Tidyverse suite of packages. This won't work unless you have manually run the chunk above, once.

```{r}
#| label = "libraries",
#| warning = FALSE
library(tidyverse)
```

## The Gapminder data

Here we load a data file from the `data/` folder inside this project.

```{r}
#| label = "load-gapminder-data"
gapminder <- read_csv(here::here("data", "gapminder.csv"))
```

Now we can look at it:

```{r}
gapminder
```

## A very first plot

We'll write our code to make a plot here. 

```{r}

```

