---
title: "Problem Set 01"
author: "<YOUR NAME HERE>"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
#| label = "setup",
#| include = FALSE
knitr::opts_chunk$set(echo = TRUE)
```

## Quarto and R Markdown

This is an Quarto document. It is a structured plain-text format that allows you to mix text and pieces (also called "chunks" or "cells") of code in a way that can be turned into HTML, PDF, and MS Word documents. For more details on using Quarto see <https://quarto.org/docs/get-started/hello/rstudio.html>.

When you click the **Render** button in RStudio a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
#| label = "addition"
1 + 1
```

## Installing Support Packages

We will only need to run this chunk _once_, so this code chunk has a flag or option that sets it not to run when the document is rendered. You must run the code manually by clicking the green "Run Current Chunk" triangle in the top right hand corner of the chunk.

```{r}
#| label = "packages",
#| eval = FALSE

## Things we'll use in the course
course_packages <- c("tidyverse", "babynames", "broom",
    "drat", "gapminder", "here", "janitor", "naniar", 
    "palmerpenguins", "skimr", "slider", "socviz",
    "usethis", "visdat","reprex", "remotes")

install.packages(course_packages, repos = "http://cran.rstudio.com")

## Data we will use
data_packages <- c("covdata", "congress", "nycdogs", 
                   "ukelection2019", "uscenpops")

remotes::install_github(paste0("kjhealy/",data_packages))
```


## The Tidyverse Packages

In this next chunk we will load the Tidyverse suite of packages. This won't work unless you have manually run the chunk above, once.

```{r}
#| label = "libraries",
#| warning = FALSE
library(tidyverse)
```

## The Gapminder data

In the next chunk we load a data file from the `data/` folder inside this project.

```{r}
#| label = "load-gapminder-data"
gapminder <- read_csv(here::here("data", "gapminder.csv"))
```

Now we can look at it:

```{r}
gapminder
```

## A very first plot

We'll write our code to make a plot here. 

```{r}

```

