---
title: "Example 03: Some basic plots"
---


We make plots with the `ggplot()` function and some `geom_` function. The former sets up the plot by specifying the data we are using and the relationships we want to see. The latter draws a specific kind of plot based on that information. 

As always we load the packages we need. Here in addition to the tidyverse we load the gapminder data as a package.

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(gapminder)
```


```{r}
gapminder
```


We saw that we can put plots in to objects, like this:

```{r}
p <- ggplot(data = gapminder, 
       mapping = aes(x = gdpPercap, y = lifeExp)) +
  geom_point()

p
```

We can set them up in a pipeline, too:

```{r}
gapminder |> 
  ggplot(mapping = aes(x = gdpPercap, y = lifeExp)) + 
  geom_point()
```


Experiment with some aesthetics, geoms, and transformations:

```{r}
gapminder |> 
  ggplot(mapping = aes(x = log(gdpPercap), y = lifeExp)) + 
  geom_point() + 
  geom_smooth()
```

```{r}
gapminder |> 
  ggplot(mapping = aes(x = lifeExp)) + 
  geom_histogram()
```


Selecting a bin width: 

```{r}
gapminder |> 
  ggplot(mapping = aes(x = lifeExp)) + 
  geom_histogram(binwidth = 5)
```




