---
title: "Example 07: Social Data and the State"
---


## Setup

```{r}

library(here)      # manage file paths
library(socviz)    # data and some useful functions
library(tidyverse) # your friend and mine

library(tidycensus) # Tidily talk to the Census
library(sf)        # Draw maps with ggplot

# Don't needlessly download geo files multiple times
options(tigris_use_cache = TRUE)

```



## Make sure you have a Census API key

As we discussed in class, you'll need an API key to talk to the Census. Get one here:

<https://api.census.gov/data/key_signup.html>

Follow the instructions and confirm the key via email when you get it. Keep the email. 

Load your key with the `census_api_key()` function. Paste your key (in quotes) instead of "YOUR API KEY GOES HERE" in the chunk below.

```{r}
# census_api_key("YOUR API KEY GOES HERE")
```

## Median Age by State

```{r}
age10 <- get_decennial(geography = "state",
                       variables = "P013001",
                       year = 2010)

age10
```

## Decennial Census Variables

There are a lot of them. 

```{r}
census_vars <- load_variables(year = 2010,
                              dataset = "sf1",
                              cache = TRUE)

census_vars
```

## Some Core Population Measures by County for North Carolina

```{r}
# Census variable names
popvars <- c("P005003", "P005004", "P005006", "P004003")


# Get a county-level dataset for NC with these variables
nc <- get_decennial(geography = "county",
                    variables = popvars,
                    year = 2010,
                    summary_var = "P001001",
                    state = "NC") |>
  mutate(pct = 100 * (value / summary_value))

nc
```

## ACS Data

```{r}
# Get median HH income by county for NC
nc_inc <- get_acs(geography = "county",
        variables = c(medincome = "B19013_001"),
        state = "NC",
        year = 2020)

nc_inc
```

With geographical information:

```{r, message=FALSE, results='hide'}
nc_inc <- get_acs(geography = "county",
                  variables = c(medincome = "B19013_001"),
                  state = "NC",
                  year = 2020,
                  geometry = TRUE)
```

```{r}
# Now we have a GEOMETRY column and spatial features information
nc_inc
```


## Draw a Map

```{r, fig.height=5, fig.width=9}
nc_inc |>
  mutate(slab = stringr::str_remove(NAME, " County.*"),
         estimate = estimate/1000) |>
  ggplot(mapping = aes(fill = estimate)) +
  geom_sf(color = "white") +
  geom_sf_text(aes(label = slab), color = "white", size = rel(0.9)) +
  scale_fill_continuous(labels = scales::label_dollar()) +
  guides(fill = guide_legend(keywidth = rel(2.2),
                             title.position = "top",
                             label.position = "bottom")) +
  labs(fill = "Median HH Income ('000s)") + 
  theme(legend.position = "bottom")

```

